---
title: "Lab 1 - Data visualization"
author: Alex LaTrenta
format: pdf
editor: visual
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
```

## Exercise 1

(Type your answer to Exercise 1 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

Problem 1:

```{r}

ggplot(data = midwest,
       aes(x = popdensity)) +
geom_histogram(binwidth = 10000) +
 labs(x = "Population Density", 
       y = "# of Midwestern Counties", 
       title = "The Population Density in Midwestern Counties")

```

The graph is very right-skewed with most states having a population density significantly lower than 25,000 in the low thousands. There are some outliers with medium and large population density.

## Exercise 2

```{r}
ggplot(data = midwest,
       aes(x = percollege, y = percbelowpoverty, color = state)) +
geom_point(alpha = .5) +
  scale_color_viridis_d() +
 labs(x = "% of People with College Degrees", 
       y = "% of People Below Poverty", 
       title = "The Poverty Line vs. College Degrees")
```


## Exercise 3

(Type your answer to Exercise 3 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

## Exercise 4

## Exercise 5

## Exercise 6

## Exercise 7
